---
title: "PML Report - Machine Learning Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prediction Assignment Write-Up


#Background and Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


#Data

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

Initialization and reading in data

```{r Initalise}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
set.seed(12345)

trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"

testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```



```{r cleaning}

training.df <- read.csv(url(trainUrl))
testcases.df <- read.csv(url(testUrl))
features <- names(testcases.df[,colSums(is.na(testcases.df))== 0])[8:59]

training.df <- training.df[,c(features,"classe")]
testcases.df <- testcases.df[,c(features,"problem_id")]

dim(training.df)
dim(testcases.df)

```

partition into training and test sets
```{r split}

inTrain <- createDataPartition(y=training.df$classe, p=0.6, list=FALSE)
myTraining <- training.df[inTrain, ]; myTesting <- training.df[-inTrain, ]
dim(myTraining); dim(myTesting)

```

Prediction algorithm - decision tree

```{r predict_algo_tree}

modFitA1 <- rpart(classe ~ ., data=myTesting, method="class")
fancyRpartPlot(modFitA1)

```


```{r predict_tree}
predictionsA1 <- predict(modFitA1, myTesting, type = "class")
confusionMatrix(predictionsA1, myTesting$classe)

```


Prediction algorithm - random forests

```{r predict_algo_rf}

modFitB1 <- randomForest(classe ~. , data=myTesting)

```


```{r predict_rf}

predictionsB1 <- predict(modFitB1, myTesting, type = "class")
confusionMatrix(predictionsB1, myTesting$classe)

```



# Generating Files 

Finally, using the provided Test Set out-of-sample error.

For Random Forests we use the following formula, which yielded a much better prediction in in-sample:

```{r generate}
predictionsB2 <- predict(modFitB1, testcases.df, type = "class")
```


```{r generate_files}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictionsB2)


```